---
title: "Advanced Computing - Data wrangling and Plotting"
author: "Brock"
date: "9/11/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r packages}
library(tidyverse) #Tidy functionality
library(dplyr) #dplyr functionality (piping)
library(here) #Helps clean pathing
library(knitr) #Rmd output
```

#Data wrangling
Just importing some data :)

```{r Data wrangling}
pubs_path <- "D:/UChicago/BSD-QBio_2020/BSD-QBio6/tutorials/advanced_computing/data/UC_Nat_Sci_1999-2019.csv"

pubs <- read.csv(pubs_path) #Reads publication data
```

#A new data type
Checking out tibbles :)

```{r A new data type}
is.data.frame(pubs) #True

pubs <- read_csv(pubs_path) #Overreads publication data as tibble. Yay!

head(pubs) #Shows first few rows
tail(pubs) #Shows last few rows
glimpse(pubs) #Shows summary of data (kinda like str in base R)
view(pubs) #See spreadsheet-like window
```

#Selecting rows and columns
Filtering data :)

```{r Selecting rows and columns}
filter(pubs, Year > 2009) %>% glimpse() #Shows publications from 2010 onward

filter(pubs, Year == 2008, 'Source title' == "Nature", 'Cited by' > 100) %>% glimpse() #Publication in Nature from 2008 with more than 100 citations

select(pubs, Authors, Title) %>% glimpse() #Shows only Authors and Title column

distinct(select(pubs,Year)) %>% glimpse() #Shows distinct years
```

#Creating pipelines using %>%
Pipelines are a great way to improve readability :)

```{r Creating Pipelines %>%}
pubs %>%
  select(Year) %>%
  distinct() %>% glimpse() #Selects distinct years, and glimpses the tibble
```

#Producing summaries
Calculate statistics using summarise :)
Statistics include: first, last, min, max, n [count the number of rows], n_distinct [count the number of distinct rows], mean, median, var, sd, etc..

```{r Producing summaries}
pubs %>% summarise(avg = mean(`Cited by`)) # 294.8619

pubs %>%  summarise(avg = mean(`Cited by`),
                    sd = sd(`Cited by`),
                    median = median(`Cited by`)) #sd = 589.0192, median = 137
```

#Summaries by group
Calculate statistics of certain groups :)

```{r Summaries by group}
pubs %>%
  group_by(`Source title`, Year) %>%
  summarise(avg = mean(`Cited by`))

#Excercise: count the number of articles by UofC researcher in Nature and Science by Source title and Year

pubs %>%
  group_by(`Source title`, Year) %>%
  summarise("Total UofC Articles" = n())
```

#Ordering the data
Use arrange(...) or arrange(desc(...)) :)

```{r Ordering the data}
pubs %>% select(Title, `Cited by`) %>% arrange(`Cited by`) %>% glimpse() #Low to high top to bottom

pubs %>% select(Title, `Cited by`) %>% arrange(desc(`Cited by`)) %>% glimpse() #High to low top to bottom
```

#Renaming columns
Use rename(...) to rename columns :)

```{r Renaming columns}
pubs %>% rename(Cites = `Cited by`) #Will not overwrite

pubs <- pubs %>% rename(Citations = `Cited by`, Journal = `Source title`) #Overwrites data
```

#Adding new variables using mutate
Use mutate() to add one or more new columns

```{r Mutate}
pubs <- pubs %>% mutate(Num_authors = str_count(Authors, ",") + 1) %>% glimpse()

pubs %>%
  group_by(Year) %>% #group papers by year
  mutate(pr = percent_rank(Citations)) %>% #compute % rank by Citations
  ungroup() %>% #remove group information
  arrange(Year, desc(pr), Authors) %>% #order by year then % rank (decreasing)
  head(20)
```

#Data plotting
Visualize your data :)

```{r Basing plotting}

```